Exploring the Role of Small Molecules in Biological Systems Using Network Approaches

### 2015-01-19 paper.tex
Slight changes

\section{The Role of Networks in Drug Discovery}
\label{sec:role-networks-drug}

>>>>

\section{The Role of Networks in Drug Discovery}
\label{sec:role-networks-drug}
Omics technologies are enabling us today to elucidate and interrogate complex relationships between proteins, genetic material and small molecules at an unprecedented speed and scale. Interrogation of relationships reveal common mechanism of action. This may result in a therapeutic candidate finding application in a seemingly unrelated disease that share a previously unknown mechanism of action. This type of relationship elucidation may also prevent unwanted side-effects if interrogation would reveal an unexpected link to a vital pathway that ought to remain unperturbed. The process of interrogation starts with experiments aimed at elucidating molecular mechanisms and pathways such as yeast two-hybrid techniques, protein and small molecule mass spectrometry, DNA microarrays and copy number variation studies. On the chemistry side, experiments may include high throughput screening of millions of chemical compounds against biological targets of therapeutic interest. These experiments result in pairwise interaction relationships from which the functional big picture biological network can then be elucidated \cite{Wu2010}. Community structure algorithm \cite{girvan2002community}, for example, allows one to discover a set of sparsely interconnected subsets of vertices, the subsets in turn however being densely connected. In the biological context, this allows connecting a set of biological cycles and smaller pathways into a big picture biological network from which inferences on new-disease application and cross-reactivity profile of a drug candidate can be drawn.
